LME-MPM applied to quasi-brittle fracture.

### 2020-10-06 2020_EFM_MPM_Eigensoftening-MacBook Pro de Pedro.tex
Revisión MMP

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init glossaries
\usepackage[acronym]{glossaries}
\makeglossaries

@@

\journal{Engineering Fracture Mechanics}
% \journal{Computers and Structures}

@@

%\title{About the modelling of fracture in quasi-brittle materials thorough the local maximum entropy material point method : the eigensoftening algorithm as proposed approach.}

\title{\textcolor{blue}{Local maximum entropy material point method applied to quasi-brittle fracture.} }

@@

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for such a methodology.  The good fitting in the predictions made by
   the eigensoftening algorithm against both analytical and experimental
  results \textcolor{red}{proves} the excellent performance of the method when challenging
  applications are to modeled.
\end{abstract}

@@

\printglossary[type=\acronymtype]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However, the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical
\textcolor{red}{artifices}  have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, or to the gradual insertion of
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries. To alleviate these shortcomings of the
\acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018}, the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}, and more
recently improving it capability to predict the crack path in 3D cases
\cite{YUN_2019}.

@@

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}

\textcolor{blue}{The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method where the continuum \gls{domain} is discretised with a finite set of material points (also known as particles in this research) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point.  In addition to this set of material points, a set of fixed background nodes is introduced to compute the balance of momentum equation. This set of nodes is defined as  $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. In the present research, any nodal quantity is denoted by ($\square_I$).}

Since the \acrshort{mpm} possesses the advantages of both Lagrangian and Eulerian descriptions, no  element distortion takes place in the \acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
\textcolor{blue}{the explicit predictor-corrector time integration scheme proposed by Molinos~\etal~\cite{Molinos2020} is adopted}. The
purpose of this choice is motivated due to its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor \textcolor{red}{$ \vec{v}_{I}^{pred}$} is computed through
\textcolor{red}{particles} velocities \textcolor{red}{$\vec{v}_p$} and accelerations \textcolor{red}{$\vec{a}_p$} and evaluated
in the $k$ time step,
%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
 \label{eq:Predictor-velocity}
    \vec{v}_{I}^{pred} =  \textcolor{blue}{\sum_{p\ \in\ \mathcal{C}}} \frac{ N_{I}(\vec{x}_p^k )\ m_p\ (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%
Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and \textcolor{red}{$N_{I}(\vec{x}_p^k) $} is the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}}
          \frac{N_{I}(\vec{x}_p^k)\ \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
        \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n +  \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
         N_{I}(\vec{x}_p^k)\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. The
  evaluation of both approaches \cite{Navas_2017_ES} against
  experimental measures show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber \textcolor{red}{reinforced} concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

\textcolor{blue}{The proposed approach to overcome the two main shortcomings of
EighenMPM  has been evaluated with three benchmark tests. On the one hand, the benchmark proposed in section \ref{sec:3.1} discuss the presence of stress instabilities provoked by the spatial discretisation. For that purpose, two different interpolation techniques, \acrshort{gimp} and \acrshort{lme}, as been compared by carrying out an eigenerosion simulation. And on the other hand the inability of the eigenerosion to simulate properly quasi-brittle fracture is discussed in sections \ref{sec:3.2} and \ref{sec:3.3}, where experimental results are compared with numerical ones. In the first case comparing with a semicircular bending test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation and \acrshort{ugimp}
\cite{Bardenhagen2004}. The application
consists of a square plate of size $H = 1$ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001. The absence of units is due to the current
simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

@@

\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 for a imposed displacement 0.015 which agrees with the

@@

  (pictures in the right side).  \textcolor{blue}{The results are printed directly on the material points.}}

@@

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\  \textcolor{red}{GPa}       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ \textcolor{red}{$\text{g/cm}^3$} \\
    Tensile strength (\gls{ft}) & 0.6\ \textcolor{red}{MPa}       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ \textcolor{red}{mm} \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocity of the top boundary particles is $v_y = 2 mm/s$, 
whereas the velocity of the bottom boundary particles is $v_y = 0 mm/s$.
To reduce dynamic effects due to a sudden loading, a soft loading ramp has been employed as a loading curve. 
Finally, the model was calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

Figure \ref{fig:Figure-Angle-Forces-test-damage-pattern} shows the
comparison of mixed-mode fracture envelope between experiments and
\acrshort{mpm} simulations for different notch inclination angles
$\alpha$ = 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. The final
fracture pattern is a vertical straight line when $\alpha$ = 0$^{\circ}$  (pure mode I). 
When $\alpha > 0^{\circ}$, the sample is
subjected to a mixed-mode pattern. For the cases of 0$^{\circ}$ and 30$^{\circ}$, cracks
initiate from the notch tip and then propagate towards upper loading
point. Experiments show that a few cracks tend to initiate behind
notch tip when the notch angle is beyond 50$^{\circ}$. Figure
\ref{fig:Figure-Angle-Forces-test-damage-pattern} demonstrates that
eigensoftening, combined with \acrshort{mpm}, is able to reproduce this
behavior. Therefore, the simulation results predict successfully the
fracture envelope in experiments and demonstrate the capability of
the eigensoftening-\acrshort{mpm} model for capturing the fracture
development under various complex loading conditions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[$\alpha$ = 0]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-0-damage-pattern.png}}
\subfigure[$\alpha$ = 30]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-30-damage-pattern.png}}
\subfigure[$\alpha$ = 60]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-60-damage-pattern.png}}
\caption{Crack pattern for different notch inclination angle, and
  comparison with the experimental results  \cite{LIM_1993}. \textcolor{blue}{The results are printed directly on the material points.}}
\label{fig:Figure-Angle-Forces-test-damage-pattern}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, \ref{fig:plot-Angle-Forces-test} compare the peak load
between experiments and simulations for the notch inclination angles
0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. This figure proofs how the
eigensoftening-\acrshort{mpm} simulation can appropriately capture the
increasing trend of the peak load with the increasing notch angles in
the experiment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{./Figure-Angle-Forces-test}
  \caption{Numerical results compared with the experimental results \cite{LIM_1993}.}
  \label{fig:plot-Angle-Forces-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init glossaries
\usepackage[acronym]{glossaries}

\makeglossaries

@@

% \journal{Engineering Fracture Mechanics}
\journal{Computers and Structures}

@@

\title{LME-MPM applied to quasi-brittle fracture.}

@@

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for such a methodology.
  %\acrshort{mpm}. 
  The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the excellent performance of the method when challenging
  applications are to modeled.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However, the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical
artifacts have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, or to the gradual insertion of
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries. To alleviate these shortcomings of the
\acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018}, the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}, and more
recently improving it capability to predict the crack path in 3D cases
\cite{YUN_2019}.

@@

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian
method where moving material points, henceforth in this research,
particles ($\square_p$) will carry on all the physical information of the local state ($\sigma_p,
\varepsilon_p$) and a set of fixed background nodes ($\square_I$) will
be introduced to compute the balance of momentum equation. Since the
\acrshort{mpm} possesses the advantages of both Lagrangian and
Eulerian descriptions, no element distortion takes place in the
\acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
an explicit predictor-corrector time integration scheme is adopted. The
purpose of this choice is motivated due to its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor $ \vec{\tilde{v}}_I^{k+1}$ is computed through
nodal velocities $\vec{v}_I$ and accelerations $\vec{a}_I$ and evaluated
in the $k$ time step,
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{\tens{m}_I^{k+1}}\ .
\end{equation}
Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and $N_{Ip}^{k}$ is
the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} =
          \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. The
  evaluation of both approaches \cite{Navas_2017_ES} against
  experimental measures show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

The proposed approach to overcome the two main shortcomings of
EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
benchmark tests''.
The first one deals with the presence of stress
instabilities provoked by the spatial discretisation, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Sections \ref{sec:3.2} and \ref{sec:3.3}, where experimental results
are compared with numerical ones, being the eigensoftening technique the one employed in this research for the
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation and \acrshort{ugimp}
\cite{Bardenhagen2004}. The application
consists of a square plate of size $H = 1\ [L] $ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current
simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025 $[L]$. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

@@

\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the

@@

  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}

@@

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocity of the top boundary particles is 
$v_y$ = − 2 mm/s, whereas the velocity of the bottom boundary particles is
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, a soft loading ramp has been employed as a loading curve. Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

Figure \ref{fig:Figure-Angle-Forces-test-damage-pattern} shows the
comparison of mixed-mode fracture envelope between experiments and
\acrshort{mpm} simulations for different notch inclination angles
$\alpha$ = 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. The final
fracture pattern is a vertical straight line when $\alpha$ = 0$^{\circ}$  (pure mode I). 
When $\alpha > 0^{\circ}$, the sample is
subjected to a mixed-mode pattern. For the cases of 0$^{\circ}$ and 30$^{\circ}$, cracks
initiate from the notch tip and then propagate towards upper loading
point. Experiments show that a few cracks tend to initiate behind
notch tip when the notch angle is beyond 50$^{\circ}$. Figure
\ref{fig:Figure-Angle-Forces-test-damage-pattern} demonstrates that
eigensoftening, combined with \acrshort{mpm}, is able to reproduce this
behavior. Therefore, the simulation results predict successfully the
fracture envelope in experiments and demonstrate the capability of
the eigensoftening-\acrshort{mpm} model for capturing the fracture
development under various complex loading conditions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[$\alpha$ = 0]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-0-damage-pattern.png}}
\subfigure[$\alpha$ = 30]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-30-damage-pattern.png}}
\subfigure[$\alpha$ = 60]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-60-damage-pattern.png}}
\caption{Crack pattern for different notch inclination angle, and
  comparison with the experimental results  \cite{LIM_1993}.}
\label{fig:Figure-Angle-Forces-test-damage-pattern}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, \ref{fig:plot-Angle-Forces-test} compare the peak load
between experiments and simulations for the notch inclination angles
0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. This figure proofs how the
eigensoftening-\acrshort{mpm} simulation can appropriately capture the
increasing trend of the peak load with the increasing notch angles in
the experiment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{./Figure-Angle-Forces-test}
  \caption{Numerical results compared with the experimental results \cite{LIM_1993}.}
  \label{fig:plot-Angle-Forces-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
